---
title: "Managing JATOS API Credentials"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Managing JATOS API Credentials}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`jatosr` never takes a token as a function argument in normal use. Your API token goes into the credential store of your operating system, the server's URL goes into a small configuration file, and every function reaches both through `jatos_connection()`. Several accounts or servers are kept apart as named profiles. This page explains where a token is kept, in which order the package looks for one, and why nothing the package writes can ever contain it.

None of the chunks are evaluated when the package is built; the output was produced against the package's mocked test server with a fake token.

## Create a token in JATOS

In the JATOS web interface open the user menu, choose *API tokens*, and create a personal access token. Give it a name that says where it will be used (`analysis-laptop`, `ci-runner`) and an expiry if the server offers one. JATOS shows the token once; it starts with `jap_`.

A JATOS token carries no roles and no scopes: it can do whatever the account it belongs to can do. For anything that runs unattended, make a dedicated API account, add it only to the studies it needs, and give its token an expiry.

## Store it once

```{r}
library(jatosr)
jatos_set_credentials("https://jatos.example.org")
```

With no `token` argument the function prompts for it, with hidden input, so the token is never typed into a script or a console history. It then:

```
#> v Stored the token of profile "default" in the "macos" credential store.
#> i The host went to
#>   '~/Library/Application Support/org.R-project.R/R/jatosr/profiles.json'.
#> v Token "analysis-laptop" for "researcher@example.org" accepted by
#>   <https://jatos.example.org> (no expiry).
```

Two different places, on purpose:

* **The token** goes into the credential store of the operating system, through the [keyring](https://keyring.r-lib.org/) package: the macOS keychain, the Windows credential store, the Secret Service on Linux. It is encrypted at rest and the package never writes it anywhere else.
* **The host** is not a secret, so it goes into `profiles.json` under `tools::R_user_dir("jatosr", "config")`, the location R sanctions for user configuration. You can read it, and `jatos_list_profiles()` can read it without touching the credential store at all.

The token is not kept in a file such as `~/.Renviron`, because a file like that is plain text: it is picked up by OneDrive, iCloud and Dropbox, it travels in backups and in copied home directories, and a project-level one reaches git more often than anyone intends.

Where `tools::R_user_dir()` is not a good place for the configuration file — a container whose home directory is read-only, a machine several people log in to — set the environment variable `JATOSR_CONFIG_DIR` to another directory. Every function of the package then reads and writes `profiles.json` there. It holds profile names and hosts only, so the directory needs to be writable, and private only as far as your hosts are.

The last line is a check against the server, once, so that a token that was pasted wrong fails here rather than three functions later. `check = FALSE` skips it; a failed check is a warning, and the token is stored anyway, so you can set credentials while the server is down.

## Where the package looks for a token

In this order, first hit wins:

1. an explicit `token =` argument to `jatos_connection()`;
2. the environment variable `JATOS_TOKEN`, or `JATOS_TOKEN_<PROFILE>` for a named profile;
3. a token already read in this session;
4. the credential store;
5. a prompt, if the session is interactive.

The host is resolved the same way, without the last two steps: `JATOS_HOST` (or `JATOS_HOST_<PROFILE>`), then `profiles.json`.

`jatos_credentials_sitrep()` says which of these each profile is actually using, without printing a token:

```{r}
jatos_credentials_sitrep()
```

```
#> -- jatosr credentials ----------------------------------------------------
#> * Packages: jatosr 0.1.0, keyring 1.4.1
#> * R version 4.6.1 (2026-06-24)
#> * Operating system: macOS Tahoe 26.6.2
#> * Credential store: "macos"
#> * Profile configuration:
#>   '~/Library/Application Support/org.R-project.R/R/jatosr/profiles.json'
#>
#> -- Profile "default" (active) --
#>
#> * Host: <https://jatos.example.org> (from config)
#> * Token: from the credential store
#>
#> -- Profile "lab_admin" --
#>
#> * Host: <https://jatos.example.org> (from config)
#> * Token: from the credential store
```

The first lines name the versions and, for a package installed from GitHub, the commit, so the report can go into a bug report as it is. Replace the host first if your server is not public.

## Continuous integration, containers and clusters

Step 2 is there for machines that have no credential store, and it is a supported path rather than a fallback. On GitHub Actions, on a cluster node, or in a Docker image, put the token in the platform's own secret store and let it arrive as an environment variable:

```yaml
env:
  JATOS_HOST: https://jatos.example.org
  JATOS_TOKEN: ${{ secrets.JATOS_TOKEN }}
```

The package reads it and needs nothing else. `jatos_set_credentials()` is not called there at all — it would refuse, because there is nowhere durable to store anything.

Because an environment variable wins over the credential store, a variable left over on your own machine will quietly shadow a token you have just stored. The package warns once per session when that happens, and `jatos_credentials_sitrep()` marks it.

## Headless machines

jatosr does not pick a credential store itself; it uses the one `keyring::default_backend()` returns. That is the Windows credential store on Windows and the keychain on macOS. On Linux it is the Secret Service (GNOME Keyring, KWallet) when two things hold: `keyring` was built with libsecret, and a desktop session provides the service. `keyring` lists libsecret as an optional system requirement: `libsecret-1-dev` on Debian and Ubuntu, `libsecret-devel` on Fedora. Install it before installing `keyring` from source, or reinstall `keyring` afterwards. Setting `R_KEYRING_BACKEND` chooses a backend by hand.

Where no system keyring exists, as on a Linux server without a desktop session, `keyring` falls back to keeping secrets in environment variables for the life of the session. Storing there would look as though it had worked and be gone at the next start of R, so `jatos_set_credentials()` refuses:

```
#> Error in `jatos_set_credentials()`:
#> ! No persistent credential store is available on this system.
#> i Set `JATOS_TOKEN` through your platform instead, on a server or in CI.
#> i Or create an encrypted file keyring once with
#>   `keyring::backend_file$new()$keyring_create("system")`.
```

Both ways out are real. The environment variable is the simpler one. The second command makes an encrypted file keyring, unlocked by a password you choose. From then on `keyring` selects the file backend by itself (from `keyring` 1.4.0 on it does so only once that keyring exists), and the package works normally. A locked file keyring still lists its entries, but a token cannot be read from it without the password, so a script unlocks it first with `keyring::keyring_unlock("system")`. Plain `keyring::keyring_create()` does not work here: it runs on the environment-variable backend, which has no keyrings.

If the store refuses the token, for example a Secret Service that stays locked, the profile's host is taken out of the configuration again, so a failed call leaves no profile with a host and no token.

## What the package checks

```{r}
jatos_has_credentials()
#> [1] TRUE
```

`TRUE` means a host and a token can be found, not that the token works. It retrieves nothing and asks the server nothing. For the real answer, `jatos_token_info()` calls the server.

Without credentials the error says which variable was looked for and what to run:

```
#> Error in `jatos_connection()`:
#> ! No `host` supplied and the `JATOS_HOST` environment variable is not set.
#> i Run `jatosr::jatos_set_credentials()` once to store it.
#> i Or set the `JATOS_HOST` environment variable.
```

A token that does not start with `jap_` is stored, with a warning, because the most common mistake is to copy the token's *name* out of the JATOS table instead of the token:

```
#> Warning message:
#> JATOS personal access tokens usually start with `jap_`.
#> i Check that you copied the token, not its name.
```

## The connection holds no token

```{r}
conn <- jatos_connection()
conn
```

```
#> <jatos_connection>
#>   profile: default
#>   host:    https://jatos.example.org
#>   api:     https://jatos.example.org/jatos/api/v1
#>   token:   from the credential store
```

There is no token in that object to print. `str()` shows everything it holds:

```{r}
str(conn)
```

```
#> List of 6
#>  $ profile   : chr "default"
#>  $ host      : chr "https://jatos.example.org"
#>  $ api_url   : chr "https://jatos.example.org/jatos/api/v1"
#>  $ id        : chr "cn_84799_2"
#>  $ auth_from : chr "keyring"
#>  $ user_agent: chr "jatosr/0.1.0"
#>  - attr(*, "class")= chr "jatos_connection"
```

`id` is an opaque label for this session; `auth_from` says which of the five steps above supplied the token. The token itself stays in an environment the package owns, for this session only, and is fetched inside the one function that builds an HTTP request.

This is what makes a connection safe to keep:

```{r}
saveRDS(conn, "conn.rds")
# the file holds the profile name and the host, and no token
```

`saveRDS()`, `save.image()`, an `.RData` written at the end of a session, a knitr cache, a targets store — none of them can write a token this way, because there is none in the object. A connection read back in another session carries a dead `id`, looks its profile up again, and works if that profile is still stored.

The package's test suite asserts this rather than trusting it: every file the package writes — the cache, the data file, the metadata sidecar, the provenance record, exported archives, `profiles.json` — is scanned byte by byte for a token, as is every message, warning and error raised during a full run.

The guarantee is about this package's objects. A token you assign to a variable of your own is an ordinary string, and is saved like one.

## Expiring tokens

`jatos_token_info()` reports what the server knows about the token in use:

```{r}
jatos_token_info()[, c("name", "expires", "expired")]
```

```
#> # A tibble: 1 x 3
#>   name            expires             expired
#>   <chr>           <dttm>              <lgl>
#> 1 analysis-laptop 2025-08-30 12:00:00 FALSE
```

Once that call has seen an expiry within seven days, the next `jatos_connection()` of the session warns, once:

```
#> Warning message:
#> The token of profile "default" expired on 2025-08-30 12:00 UTC.
#> i Create a new token in JATOS and store it with `jatos_set_credentials()`.
```

Servers that never expire tokens report no expiry at all, and nothing is warned about.

## Several accounts or servers: profiles

A profile is a name for one host-and-token pair. The default profile is called `default` and needs no name anywhere.

```{r}
jatos_set_credentials("https://jatos.example.org", profile = "lab_admin")
```

```
#> v Stored the token of profile "lab_admin" in the "macos" credential store.
#> i The host went to
#>   '~/Library/Application Support/org.R-project.R/R/jatosr/profiles.json'.
#> i Use it with `jatos_connection("lab_admin")` or set `JATOS_PROFILE` to
#>   "lab_admin".
```

The profile name is the username of the credential store entry, so profiles and entries map one to one and you can see them in Keychain Access or seahorse under the service `jatosr`.

Use one either per call or for the whole session:

```{r}
admin <- jatos_connection("lab_admin")
jatos_studies(conn = admin)

Sys.setenv(JATOS_PROFILE = "lab_admin")   # or set it in ~/.Renviron
jatos_studies()                           # now uses lab_admin
```

`jatos_list_profiles()` shows what exists, from all three places, and never a token:

```{r}
jatos_list_profiles()
```

```
#> # A tibble: 3 x 5
#>   profile   host                        has_token active auth_from
#>   <chr>     <chr>                       <lgl>     <lgl>  <chr>
#> 1 default   https://jatos.example.org   TRUE      TRUE   keyring
#> 2 lab_admin https://jatos.example.org   TRUE      FALSE  keyring
#> 3 other_lab https://jatos.other-lab.org TRUE      FALSE  keyring
```

## Removing a profile

```{r}
jatos_remove_credentials(profile = "other_lab")
```

```
#> ! About to remove the credentials of profile "other_lab" from the "macos"
#>   credential store.
#> i A token that is stored nowhere else cannot be recovered; JATOS shows it
#>   once.
#>
#> 1: Yes, remove them
#> 2: No, keep them
#>
#> Selection: 1
#> v Deleted the token of profile "other_lab" from the "macos" credential
#>   store.
#> v Removed its host from
#>   '~/Library/Application Support/org.R-project.R/R/jatosr/profiles.json'.
#> i The token itself is still valid; delete it in the JATOS user menu under
#>   API tokens to revoke it.
```

Pass `confirm = FALSE` in a script. Calling it again says there was nothing to do rather than failing:

```
#> i Nothing to remove: profile "other_lab" is not in the credential store or
#>   the configuration.
#> i `jatos_list_profiles()` shows what is set.
```

Two situations draw a warning: an environment variable still defines the profile, in which case it is still in use and will be back at the next start of R if an `.Renviron` sets it; and `JATOS_PROFILE` still names the profile you just removed, in which case every call without a `conn` fails until you unset it.

## What this protects against, and what it does not

The credential store is unlocked when you log in. The token is encrypted on disk, and it is readable by any process running as you.

* **Protected:** cloud sync, backups, shared folders, a copied home directory, an accidental commit, a screen-shared script, a workspace someone else opens.
* **Not protected:** code running in your own R session, or as your user, that goes looking for it.

If you want more than that, make a keyring of its own with a password you set and lock it when you are not using it (`keyring::keyring_create()`, `keyring::keyring_lock()`).

## Revoking

Removing a profile deletes your copy of the token. It does not revoke it: the token keeps working for anyone who has it. A token that has been pasted somewhere it should not be (a chat, a committed file, a rendered report) is revoked in the same JATOS menu that created it. Then either store a new one with `jatos_set_credentials()`, with the same `profile` if it was not the default, which replaces the old one, or drop the profile altogether with `jatos_remove_credentials()`.
